---
title: "Notes on assembling the dataset for Elise Collett, student at St.FX"
author: "Daniel Ricard"
date: '`r paste0("Last modified: ", Sys.time())`'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(gulf)
library(knitr)
library(kableExtra)
library(dplyr)
library(ggplot2)
library(worms)
```

```{r, include=FALSE}
taxo.tree.fct <- function(aphia.id, species.code){
my.df <- wormsbyid(aphia.id)

## what taxonomic rank are we dealing with?
if(my.df$rank=="Subspecies"){
  vars <- c("kingdom","phylum","class","order","family","scientificname")
  return(cbind(species.code, my.df[,vars]))
}
if(my.df$rank=="Species"){
  vars <- c("kingdom","phylum","class","order","family","scientificname")
  return(cbind(species.code, my.df[,vars]))
}
if(my.df$rank=="Subgenus"){
  vars <- c("kingdom","phylum","class","order","family","scientificname")
  return(cbind(species.code, my.df[,vars]))
}
if(my.df$rank=="Genus"){
  vars <- c("kingdom","phylum","class","order","family","scientificname")
  return(cbind(species.code, my.df[,vars]))
}
if(my.df$rank=="Superfamily"){
  vars <- c("kingdom","phylum","class","order","family","scientificname")
  return(cbind(species.code, my.df[,vars]))
}
if(my.df$rank=="Family"){
  vars <- c("kingdom","phylum","class","order","family","scientificname")
  return(cbind(species.code, my.df[,vars]))
}
if(my.df$rank=="Infraorder"){
  vars <- c("kingdom","phylum","class","order","family","scientificname")
  return(cbind(species.code, my.df[,vars]))
}
if(my.df$rank=="Order"){
  my.df$family<-NA
  vars <- c("kingdom","phylum","class","order","family","scientificname")
  return(cbind(species.code, my.df[,vars]))
}
if(my.df$rank=="Class"){
  my.df$family<-NA
  my.df$order<-NA
  vars <- c("kingdom","phylum","class","order","family","scientificname")
  return(cbind(species.code, my.df[,vars]))
}
if(my.df$rank=="Subphylum"){
  my.df$family<-NA
  my.df$order<-NA
  my.df$class<-NA
  vars <- c("kingdom","phylum","class","order","family","scientificname")
  return(cbind(species.code, my.df[,vars]))
}
if(my.df$rank=="Phylum"){
  my.df$family<-NA
  my.df$order<-NA
  my.df$class<-NA
  vars <- c("kingdom","phylum","class","order","family","scientificname")
  return(cbind(species.code, my.df[,vars]))
}


}## end function


```
## Collaboration with St. FX

### Species considered
To get things started, I am generating a dataset similar to that used by Benoit and Swain (2008). Table A1 shows 52 species. Let's try and reproduce that table by using the correct species codes in the RV database.

```{r species1, cache=TRUE, include=FALSE}
sp.codes <- c(241, 201, 204, 202, 220, 221, 712, 62, 
              60, 110, 10, 118, 11, 16, 410, 14, 114, 
              112, 12, 361, 64, 63, 610, 50, 122, 623, 
              622, 624, 616, 598, 646, 640, 41, 40, 
              30, 42, 43, 31, 143, 340, 350, 880, 314, 
              300, 316, 301, 304, 501, 503, 320, 307, 23)

## add additional taxonomic information from WoRMS
## Class - Order - Family, to follow Table A1 in Benoit & Swain
for.table <- 
  merge(
  merge(
  merge(
    data.frame(
  num=1:length(sp.codes),
  species.code=sp.codes
  ),
  species.str(sp.codes, "latin", as.data.frame=TRUE),
  by.x="species.code", by.y="CODE"),
  species.str(sp.codes, "english", as.data.frame=TRUE),
  by.x="species.code", by.y="CODE"),
  species.code(sp.codes, input.coding="RV", output.coding="worms", as.data.frame=TRUE),
  by.x="species.code", by.y="rv_code")


taxo.list.out <- lapply(1:length(sp.codes), function(ss){
  aa1<-for.table[ss,"aphia_id"];
  aa2<-for.table[ss,"species.code"];
  taxo.tree.fct(aa1,aa2)})

taxo.t <- do.call(rbind, taxo.list.out)

taxo.df.out <- merge(for.table, taxo.t, by="species.code")

taxo.df.out$species.code <- as.character(taxo.df.out$species.code)

## species that need to be summed over a number of species codes
## Seasnails, Liparis species, species codes 503, 504, 505, 506, 
## Alligatorfishes, 2 species
## Hookear sculpins, 2 species

## manually fix Liparis species
taxo.df.out[taxo.df.out$species.code=="503","LATIN"] <- "Liparis sp."
taxo.df.out[taxo.df.out$species.code=="503","ENGLISH"] <- "Seasnails"
taxo.df.out[taxo.df.out$species.code=="503","aphia_id"] <- 126160
taxo.df.out[taxo.df.out$species.code=="503","species.code"] <- "503, 504, 505, 506, 508, 512, 513"

## manually fix alligatorfishes
taxo.df.out[taxo.df.out$species.code=="340","LATIN"] <- "Aspidophoroides monopterygius, Aspidophoroides olrikii"
taxo.df.out[taxo.df.out$species.code=="340","ENGLISH"] <- "Alligatorfishes"
taxo.df.out[taxo.df.out$species.code=="340","aphia_id"] <- 159458
taxo.df.out[taxo.df.out$species.code=="340","species.code"] <- "340, 341"

## manually fix hookear scuplins
taxo.df.out[taxo.df.out$species.code=="880","LATIN"] <- "Artediellus sp."
taxo.df.out[taxo.df.out$species.code=="880","ENGLISH"] <- "Hookear sculpins"
taxo.df.out[taxo.df.out$species.code=="880","aphia_id"] <- 126147
taxo.df.out[taxo.df.out$species.code=="880","species.code"] <- "306, 880"

## manually fix eelpout species
taxo.df.out[taxo.df.out$species.code=="598","ENGLISH"] <- "Eelpouts"
taxo.df.out[taxo.df.out$species.code=="598","species.code"] <- "598,619,620,627,628,641,642,643,647"

## manually add Class Actinopteri to sea raven
taxo.df.out[taxo.df.out$LATIN=="Hemitripterus americanus","class"] <- "Actinopteri"



taxo.df.out$aphia.url <- paste("http://www.marinespecies.org/aphia.php?p=taxdetails&id=", taxo.df.out$aphia_id, sep="")

vars <- c("class","order","family","LATIN","ENGLISH","aphia_id","species.code")
o1 <- order(taxo.df.out$num)
urls <- taxo.df.out$aphia.url[o1]

csv.fn1 <- paste0("sGSL-RV-data-for-Elise-", format(Sys.time(), "%Y-%m-%d"),"-species.csv")
write.csv(taxo.df.out[o1,vars], file=csv.fn1)

```


```{r species2, results='asis', echo=FALSE}
kable(taxo.df.out[o1,vars], row.names=FALSE, longtable=TRUE, caption="The 52 species of interest, meant to match Table A1 in Benoit and Swain.") %>% 
  column_spec(4, width="2.5cm") %>% 
  column_spec(7, width="2.0cm") %>% 
  collapse_rows(columns=1:3, row_group_label_position = "stack", latex_hline="major")

```

### Dataset of abundance per tow
Now extract set cards and catch cards for representative sets in strata 415 to 439, and compute the yearly stratified random estimate of abundance per tow, corrected for gear, vessel and diurnal effects.


```{r computeCmatrix, cache=TRUE, results="hide", fig.width=6, fig.height=8}
data(rv)

yrs <- 1971:2020
x <- rv.good.sets(yrs)
x <- x[x$stratum %in% c(415:439),]
x$unique.id <- paste(x$vessel.code, x$year, x$cruise.number, x$set.number, sep="-")

y <- rv$cat
y$unique.id <- paste(y$vessel.code, y$year, y$cruise.number, y$set.number, sep="-")
y<-adjust(y,x)

## output matrix, called C in Benoit & Swain
C.matrix <- matrix(NA, nr=length(sp.codes), nc=length(yrs), dimnames=list(species=sp.codes, year=yrs))
C.df <- expand.grid(year=yrs, species=sp.codes, normalised=NA)
## same matrix, but not normalised to 0-1
C.stratified.matrix <- matrix(NA, nr=length(sp.codes), nc=length(yrs), dimnames=list(species=sp.codes, year=yrs))


for(i in 1:length(sp.codes)){
  s <- sp.codes[i]
  if(s==503) { # snailfishes
    ss <- c(503, 504, 505, 506, 508, 512, 513)
    this.y <- y[y$species %in% ss,]
    ## sum for each set
    agg.df <- aggregate(number.caught~unique.id, this.y, sum)
    vars <- c("unique.id", key(x))
    this.agg.y <- merge(x[,vars], agg.df, by="unique.id")
    this.agg.y$species<-503
    z <- merge.catch(x, rvcat(this.agg.y), var = "number.caught") #
  }
    if(s==340) { # alligatorfishes
    ss <- c(340,341)
    this.y <- y[y$species %in% ss,]
    ## sum for each set
    agg.df <- aggregate(number.caught~unique.id, this.y, sum)
    vars <- c("unique.id", key(x))
    this.agg.y <- merge(x[,vars], agg.df, by="unique.id")
    this.agg.y$species<-340
    z <- merge.catch(x, rvcat(this.agg.y), var = "number.caught")
  }
    if(s==880) { # hookear sculpins
    ss <- c(306, 880)
    this.y <- y[y$species %in% ss,]
    ## sum for each set
    agg.df <- aggregate(number.caught~unique.id, this.y, sum)
    vars <- c("unique.id", key(x))
    this.agg.y <- merge(x[,vars], agg.df, by="unique.id")
    this.agg.y$species<-880
    z <- merge.catch(x, rvcat(this.agg.y), var = "number.caught")
    }
  if(s==598) { # eelpouts
    ss <- c(598,619,620,627,628,641,642,643,647)
    this.y <- y[y$species %in% ss,]
    ## sum for each set
    agg.df <- aggregate(number.caught~unique.id, this.y, sum)
    vars <- c("unique.id", key(x))
    this.agg.y <- merge(x[,vars], agg.df, by="unique.id")
    this.agg.y$species<-598
    z <- merge.catch(x, rvcat(this.agg.y), var = "number.caught")
    }
  else{
    this.y <- y[y$species==s,]
    z <- merge.catch(x,this.y)
  }
  
stratified.number.df <- smean(z, "number.caught", by=c("year"))
matrix.row <- (stratified.number.df$mean - min(stratified.number.df$mean)) / (max(stratified.number.df$mean) - min(stratified.number.df$mean))

C.matrix[i,] <- matrix.row
C.df[C.df$species==s,"normalised"] <- matrix.row
C.stratified.matrix[i,] <- stratified.number.df$mean
}

# visualize C matrix using ggplot
C.df$species.name <- species.str(C.df$species, "english")
g <- ggplot(data=C.df,
       aes(x=year, y=species.name, fill=normalised)) + geom_tile() + 
       theme_bw() + scale_fill_gradient(low = "white", high = "black")
g

```

### Tow-level dataset

Also provide tow-level data for exploration purposes.

```{r writeCSV}
strat.stats <- stratum.info(region = "gulf", survey = "rv", stratum = 415:439)

## write to CSV file
csv.fn2 <- paste0("sGSL-RV-data-for-Elise-", format(Sys.time(), "%Y-%m-%d"),"-strata-stats.csv")
keep.vars <- c("stratum","area","trawlable.units")
write.csv(strat.stats[,keep.vars], file=csv.fn2)

csv.fn3 <- paste0("sGSL-RV-data-for-Elise-", format(Sys.time(), "%Y-%m-%d"),"-sets.csv")
x.keep.vars <- c("unique.id","year","month","day","vessel.code","cruise.number","gear","stratum","set.number","start.hour","start.minute","start.second","experiment","duration","longitude.start","latitude.start","longitude.end","latitude.end","depth.start","depth.end","distance","species.fish.number","species.invertebrate.number","catch.total.weight","surface.temperature","bottom.temperature","bottom.salinity")
o.x <- order(x$year, x$set.number)
write.csv(x[o.x,x.keep.vars], file=csv.fn3)


csv.fn4 <- paste0("sGSL-RV-data-for-Elise-", format(Sys.time(), "%Y-%m-%d"),"-catch.csv")
y.keep.vars <- c("unique.id", "species", "number.caught", "weight.caught")
k.spec.keep <- c(strsplit(paste0(taxo.df.out$species.code, collapse="", sep=","), ",")[[1]])
write.csv(y[y$species %in% k.spec.keep,y.keep.vars], file=csv.fn4)

csv.fn5 <- paste0("sGSL-RV-data-for-Elise-", format(Sys.time(), "%Y-%m-%d"),"-C-matrix.csv")

write.csv(C.matrix, file=csv.fn5)

csv.fn6 <- paste0("sGSL-RV-data-for-Elise-", format(Sys.time(), "%Y-%m-%d"),"-C-stratified-matrix.csv")
write.csv(C.stratified.matrix, file=csv.fn6)

## create a backbone Excel file to copy and paste the CSV files into
library(xlsx)
xl.fn <- "sGSL-RV-data-for-Elise-2021-10-12.xlsx"
t.df <- data.frame(x=1)
write.xlsx(t.df, file=xl.fn, sheetName = "Species list")
write.xlsx(t.df, file=xl.fn, sheetName = "Strata statisitics", append = TRUE, row.names=FALSE)
write.xlsx(t.df, file=xl.fn, sheetName = "Sets", append = TRUE, row.names=FALSE)
write.xlsx(t.df, file=xl.fn, sheetName = "Adjusted catch", append = TRUE, row.names=FALSE)
write.xlsx(t.df, file=xl.fn, sheetName = "C matrix", append = TRUE, row.names=FALSE)
write.xlsx(t.df, file=xl.fn, sheetName = "C stratified matrix", append = TRUE, row.names=FALSE)

```

The xlsx package in R was giving a memory error when writing an Excel file, so I am outputting CSV files and assembling them in an Excel file manually instead.
